\documentclass[8pt, landscape, fleqn]{scrartcl}
\setlength{\parindent}{0pt}
\usepackage[ngerman]{babel}
%\usepackage[applemac]{inputenc}
\usepackage[utf8]{inputenc}
\usepackage[dvips]{geometry}
\usepackage{latexsym}
\usepackage{multicol}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{array}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{ulem}
\usepackage{amsfonts}
\usepackage{dsfont}
\usepackage{charter} %%% Schreibart
%\renewcommand{\familydefault}{\sfdefault}



%%%%%%%%%%Paket für Chemische Formeln
\usepackage{chemformula} 
\usepackage[version=3]{mhchem}
%%%%%%%%%%%%%%%%% Farbe
\usepackage{color}

\pagestyle{plain}
\typearea{45}
\columnsep 39pt
\columnseprule .4pt
\setlength{\extrarowheight}{0.9em}

\renewcommand{\arraystretch}{0.8}

\makeatletter
\renewcommand{\section}{\@startsection{section}{1}{0mm}%
{-2\baselineskip}{0.8\baselineskip}%
{\hrule depth 0.2pt width\columnwidth\hrule depth1.5pt
width0.25\columnwidth\vspace*{1.2em}\Large\bfseries\rmfamily}}
\makeatother


\makeatletter
\renewcommand{\subsection}{\@startsection{subsection}{1}{0mm}%
{-2\baselineskip}{0.8\baselineskip}%
{\hrule depth 0.2pt width\columnwidth\hrule depth0.75pt
width0.25\columnwidth\vspace*{1.2em}\large\bfseries\rmfamily}}
\makeatother

\makeatletter
\renewcommand{\subsubsection}{\@startsection{subsubsection}{1}{0mm}%
{-2\baselineskip}{0.8\baselineskip}%
{\hrule depth 0.2pt width\columnwidth\vspace*{1.2em}\normalsize\bfseries\rmfamily}}
\makeatother

\newcommand{\Mx}[1]{\begin{bmatrix}#1\end{bmatrix}}
\begin{document}
\part*{\LARGE\textrm{Compressible Flows $\hfill$ Xeno Meienberg}}
\begin{multicols*}{3}

\section{General Considerations}

\begin{equation*}
    \frac{D\rho}{D t} = \frac{\partial \rho}{\partial t} + u_i \frac{\partial \rho}{\partial x_i} \neq 0
\end{equation*}

\begin{itemize}
    \item Wave propagation
    \item Convective flows with buoancy
    \item Flows with variable temperature, friction, sources of heat
    \item High speed flows with Mach numbers $Ma \geq 1$
\end{itemize}

Compressible flows can still be described through the continuum model and conservation laws. The assumption is also that the thermodynamic state of the fluid is in a local equilibrium. \\

\textbf{Assumptions}
\begin{itemize}
    \item Length scale of flows \emph{large} compared to molecular scales (mean free path $\lambda$)
    \item Length scale of flows \emph{small} compared to the geometric scales (length $L$)
    \item Time scale $\tau_F$ of the flow \emph{long} compared to the molecular process (relaxation) time constants $\tau_R$
\end{itemize}


\textbf{Description of the ``Continuum'' Flow State}
\begin{itemize}
    \item Three components of flow velocity $\underline{u}(\underline{x},t)$
    \item The fluid density $\rho(\underline{x},t)$
    \item The fluid pressure $p(\underline{x},t)$
    \item The energy $e(\underline{x},t)$
\end{itemize}

The required equations are the conservation laws for mass, momentum and energy together with suitable thermodynamic equations of state. With corresponding initial and boundary conditions, the evolution can then be computed.

\section{Thermodynamic Relations}

\textbf{State Variables}
\begin{itemize}
    \item Density: $\rho = \rho(p,T)$
    \item Pressure: $p = p(\rho,T)$
    \item Temperature: $T = T(\rho, p)$
    \item Internal energy: $e = e(\rho, T)$ $[e] = J/kg$
    \item Enthalpy: $h = h(p,T)$
    \item Entropy: $s = s(\rho, T)$
\end{itemize}
\textbf{Van der Waals Gas}
\begin{equation*}
    (p+a\rho^2)\left(\frac{1}{\rho}-b\right) = RT
\end{equation*}
\textbf{Incompressible Fluid}
\begin{equation*}
    \rho = const. \neq \rho(p,T)
\end{equation*}
\section{Conservation Laws for Continuum Flows}
\underline{\textbf{Mass Conservation}} \\
\begin{equation*}
	\frac{Dm}{Dt} = \frac{D}{Dt}\int_{\tilde V} \rho d\tilde V = 0 ~\text{(material~ volume)}
\end{equation*}
\begin{equation*}
	\int_V \frac{\partial \rho}{\partial t} dV + \int_S \rho(\bold{u}\cdot \bold{n}) dS = 0~\text{(Eulerian~Volume)}
\end{equation*}
\begin{equation*}
	\frac{\partial \rho}{\partial t} + \frac{\partial}{\partial x_i}(\rho u_i) = 0~\text{(material~volume~/~index)}
\end{equation*}
\begin{equation*}
	\frac{D\rho}{Dt} = -\rho \frac{\partial u_i}{\partial x_i}~\text{(Eulerian~Volume / ~index)}
\end{equation*}

\underline{\textbf{Momentum Conservation}}
\begin{equation*}
    \frac{\partial}{\partial t} (\rho u_i) + \frac{\partial}{\partial x_j} (\rho u_i u_j) = \frac{\partial}{\partial x_j} \sigma_{ij} + \rho f_i
\end{equation*}
\begin{equation*}
    \rho \frac{D u_i}{Dt} = \frac{\partial}{\partial x_j} \sigma_{ij} + \rho f_i
\end{equation*}
\begin{equation*}
    \sigma_{ij} = -p \delta_{ij} + \tau_{ij}
\end{equation*}
\begin{equation*}
    \tau_{ij} = \mu\left(\frac{\partial u_i}{ \partial x_j} + \frac{\partial u_j}{\partial x_i}\right) + \left( \mu_v - \frac{2}{3} \mu\right) \delta_{ij} \frac{\partial u_k}{\partial x_k}
\end{equation*}
\begin{equation*}
    \rho \frac{D u_i}{D t} = -\frac{\partial p}{\partial x_i} + \frac{\partial}{\partial x_j}\left[\mu \left( \frac{\partial u_i}{x_j} + \frac{\partial u_j}{\partial x_i}\right) + \left( \mu_v - \frac{2}{3}\mu\right) \delta_{ij} \frac{\partial u_k}{\partial x_k}\right] + \rho f_i
\end{equation*}

\underline{\textbf{Energy Conservation}}
\begin{equation*}
    \rho \frac{D}{Dt}(e + \frac{1}{2} u_1^2 ) = \frac{\partial}{\partial x_j} (\sigma_{ij} u_i)) + \rho f_i u_i - \frac{\partial q_i}{\partial x_i} + \rho q_v 
\end{equation*}
\begin{equation*} \rho \frac{D}{Dt}(e + \frac{1}{2} u_1^2 ) = -\frac{\partial}{\partial x_i}(pu_i) + \frac{\partial }{\partial x_j}(\tau_{ij} u_i) + \rho f_i u_i - \frac{\partial q_i}{\partial x_i} + \rho q_v
\end{equation*}
    $\rho u_i \frac{D u_i}{Dt} = \rho \frac{D}{Dt}\left(\frac{u_i^2}{2}\right) = -u_i \frac{\partial p}{\partial x_i} + u_i \frac{\partial}{\partial x_j} \tau_{ij} + \rho f_i u_i$
    $\rho \frac{D e}{D t} = \rho \frac{D}{Dt}\left(e + \frac{1}{2} u_i^2 \right) - \rho \frac{D}{Dt}\left( \frac{u_i^2}{2}\right) = \\ = -p \frac{\partial u_i}{\partial x_i} + \tau_{ij} \frac{\partial u_i}{\partial x_j} + \rho q_v - \frac{\partial q_i}{\partial x_i}$ \\

\underline{\textbf{Dissipation Function $ \Phi $}} \\
Insert $h = e + \frac{p}{\rho}$ to obtain Enthalpy equation, introduce $h_t = h + \frac{u_i^2}{2}$ and add kinetic energy (p. 15). For perfect gasses, $h = c_p T$, $q_i = -k \frac{dT}{dx}$, derive the temperature equation. \\

\textbf{Entropy Equation}
\begin{equation*}
    \rho T \frac{Ds}{Dt} = \Phi + \rho q_v - \frac{\partial q_i}{\partial x_i}
\end{equation*}
\textbf{Vorticity Equation}
\begin{equation*}
    \rho \frac{D}{Dt} \left( \frac{\vec{\omega}}{\rho} \right) = \left( \vec{\omega} \cdot \nabla \right) \vec{u} + \frac{1}{\rho^2} \nabla \rho \times \nabla p + \nabla \times \left( \frac{1}{\rho} \nabla \cdot \vec{\tau}\right)
\end{equation*}
\textbf{Crocco Theorem (rewritten momentum equation using Enthalpy and Entropy)}
\begin{equation*}
    \frac{\partial u}{\partial t} + \nabla \left( \frac{1}{2} \vec{u}^2 + h + \psi \right) = \vec{u} \times \vec{\omega} + T \nabla s + \frac{1}{\rho} \nabla \cdot \vec{\tau}
\end{equation*}

\textbf{Compressible Bernoulli}
equation (integrate momentum equation law along particle path). Clasical not feasible
\begin{align*}
    \rho \left( \frac{D h_t}{D t} - f_i u_i\right) = 0 \\
    f_i = - \frac{\partial \psi}{\partial x_i} \\
    \psi \neq \psi(t) \\
    \frac{D}{Dt}\left( h_t + \psi \right) = 0
\end{align*}
Between 2 points along stream line
\begin{equation*}
    h_t + \psi = e + \frac{p}{\rho} + \frac{u_i^2}{2} + \psi = const.
\end{equation*}

\section{Simplification Strategies (p.20)}

\begin{itemize}
    \item Unsteady $\rightarrow$ steady (no wave propagation) (no time dependence)
    \item 3D $\rightarrow$ 2D $\rightarrow$ quasi 1-D
    \item Viscous, heat conduction $\rightarrow$ inviscid, adiabatic (isentropic, homentropic)
    \item Subsonic $\rightarrow$ transonic $\rightarrow$ supersonic $\rightarrow$ hypersonic (Elliptic $\rightarrow$ hyperbolic)
    \item Full nonlinear $\rightarrow$ linearised (solve for small pertubations around predefined flow state unique solvable problem, separation of influencing factors facilitated)
\end{itemize}

\section{Conservation Laws for Stream Tubes (p. 22)}

Quasi 1D, separate for environment. Outer surface formed by instantaneous streamlines, no flow across boundaries. Inlet + outlet. Shape (t). For small enough $A$, flow properties can be treated constant in any cross section. \\

\textbf{Mass Conservation}
\begin{equation*}
    \int_1^2 \frac{\partial}{\partial t}\left[ \rho(s,t) A(s,t)\right] ds + \rho_2 A_2 u_2 - \rho_1 A_1 u_1 = 0
\end{equation*}
\begin{equation*}
    \dot{m} = \rho A u = const.
\end{equation*}

\textbf{Momentum Conservation}
\begin{align*}
    \int_1^2 \frac{\partial}{\partial t} \left[ \rho(s,t) A(s,t) \right] ds + \rho_2 A_2 u_2 \vec{u}_2 - \rho_1 A_1 u_1 \vec{u}_1 = \\
    = -p_2 A_2 \vec{n}_2 + p_1 A_1 \vec{n}_1 + F_\tau |_1^2 + F_S
\end{align*}

Steady, frictionless
\begin{equation*}
    \rho_2 u_2^2 + p_2 = \rho_1 u_1^2 + p_1
\end{equation*}

\textbf{Energy Conservation (p.20)}

Steady, frictionless
\begin{equation*}
    e_2 + \frac{u_2^2}{2} + \frac{p_2}{\rho_2} = e_1 + \frac{u_1^2}{2} + \frac{p_1}{\rho_1}
\end{equation*}

Enthalpy subsitution $h = e + \frac{p}{\rho} \rightarrow h_{t1} = h_{t2} = const.$

\section{Steady one-dimensional Flow without Friction and Heat (p. 25)}

Assumptions: 

\begin{itemize}
    \item No friction (inviscid)
    \item No heat source or transport
    \item No flow through mantle
    \item Perfect gas
\end{itemize}
\begin{equation*}
    Ma = \frac{u}{a}
\end{equation*}
\begin{equation*}
    a^2 = \gamma R T 
\end{equation*}
Stagnation properties, when $u=0$ (Ruhegrösse), subscript 0: 
\begin{equation*}
    \frac{h_0}{h} = \frac{T_0}{T} = \left( \frac{a_0^2}{a^2}\right) = 1 + \frac{\gamma - 1}{2} Ma^2
\end{equation*}

Isentropic flow (p.26):
\begin{equation*}
    \frac{p_0}{p} = \left( \frac{T_0}{T}\right)^{\frac{\gamma}{\gamma -1}} = \left[ 1 + \frac{\gamma-1}{2} Ma^2 \right]^{\frac{\gamma}{\gamma-1}}
\end{equation*}
\begin{equation*}
    \frac{\rho_0}{\rho} = \left( \frac{T_0}{T} \right)^{\frac{1}{\gamma-1}} = \left[ 1 + \frac{\gamma-1}{2} Ma^2 \right]^{\frac{1}{\gamma-1}}
\end{equation*}
When $Ma < 0.3$, density changes $< 4.5 \%$: Assumption is: incompressible. The critical state is then ($Ma = 1$), $superscript~^*$
\begin{equation*}
    \frac{h^*}{h_0} = \frac{T^*}{T_0} = \left( \frac{a^{*2}}{a_0^2}\right) = \left[ 1 + \frac{\gamma -1 }{2}\right]^{-1} = \frac{2}{\gamma+1} = 0.8333 (\gamma = 1.4)
\end{equation*}
\begin{equation*}
    \frac{p^*}{p_0} = \left( \frac{2}{\gamma+1} \right)^{\frac{\gamma}{\gamma-1}} = 0.5283 (\gamma = 1.4)
\end{equation*}
\begin{equation*}
    \frac{\rho^*}{\rho_0} = \left( \frac{2}{\gamma+1} \right)^{\frac{1}{\gamma-1}} = 0.6339 (\gamma = 1.4)
\end{equation*}
Critical $Ma^*$ (isentropic flow stays limited when $Ma \rightarrow \infty$). The flow velocity stays finite even if $Ma$ goes to infinity:
\begin{align*}
    Ma^* = \frac{u}{a^*} = \frac{u}{a(Ma=1)} = \frac{u}{a}\frac{a}{a_0}\frac{a_0}{a^*} \\
    = Ma \sqrt{\frac{T}{T_0}}\sqrt{\frac{T_0}{T^*}} = \sqrt{\frac{\frac{\gamma+1}{2}Ma^2}{1+\frac{\gamma-1}{2}Ma^2}}
\end{align*}
\begin{equation*}
    Ma^* \rightarrow \sqrt{\frac{\gamma+1}{\gamma-1}}~(Ma \rightarrow \infty) = 2.4495~(\gamma = 1.4)
\end{equation*}

\underline{\textbf{Area velocity relation}}

A velocity increase $\rightarrow$ density decrease (always). If $Ma << 1$, then the density changes are small compared to the velocity changes. A small velocity increase at $Ma >> 1$ will lead to large density changes.
\begin{equation*}
    Ma^2 \frac{1}{u} \frac{d u}{d x} = -\frac{1}{\rho} \frac{d \rho}{d x}~~\textbf{(Mach-density relation)}
\end{equation*}
\begin{equation*}
    \left( Ma^2 -1 \right) \frac{1}{u} \frac{d u}{d x} = \frac{1}{A} \frac{d A}{d x}~~\textbf{(Mach-Area relation)}
\end{equation*}

If $Ma < 1$, then an area increase will result in a velocity reduction. If $Ma > 1$, then opposite applies. If $Ma = 1$, then a change in Area $A$ has no effect (chocked flow) \\

\underline{\textbf{Stationary normal shock}}
\begin{align*}
    \frac{u_2}{u_1} &= \frac{\rho_1}{\rho_2} = 1 - \frac{2}{\gamma+1} \left( 1- \frac{1}{Ma_1^2}\right) = \frac{1}{Ma^{*2}} \\
    \frac{p_2}{p_1} &= 1 + \frac{2 \gamma}{\gamma + 1} \left( Ma_1^2 - 1 \right) \\
    \frac{T_2}{T_1} &= \left[ 1 + \frac{2 \gamma}{\gamma + 1} \left( Ma_1^2 -1 \right) \right]\left[ 1-\frac{2}{\gamma+1} \left( 1-\frac{1}{Ma_1^2} \right) \right] \\
    &\frac{\Delta s}{R} = \frac{1}{\gamma-1} \left[ \ln \left( \frac{p_2}{p_1} \right) - \gamma \ln \left( \frac{\rho_2}{\rho_1}\right)\right] = \\
    & \frac{1}{\gamma-1}\left\{\left[ 1 + \frac{2\gamma}{\gamma+1}\left( Ma_1^2 -1 \right)\right] \left[ 1-\frac{2}{\gamma+1}\left(1-\frac{1}{Ma_1^2}\right)\right]\right\}
\end{align*}

$h_{01} = h_{02}$, $T_{01} = T_{02}$, and total enthalpy conserved (however stagnation pressure not constant, $p_{01}\neq p_{02}$):
\begin{align*}
    &\frac{p_{02}}{p_{01}} = \frac{p_{02}}{p_{2}} \frac{p_2}{p_1} \frac{p_1}{p_{01}} = \frac{p_2}{p_1} \left( \frac{T_{02}}{T_2} \right)^{\frac{\gamma}{\gamma-1}}\left(\frac{T_1}{T_{01}}\right)^{\frac{\gamma}{\gamma-1}} = \\
    & \left[ 1 + \frac{2 \gamma}{\gamma + 1} (Ma_1^2 -1)\right]^{\frac{\-1}{\gamma-1}} \left[1- \frac{2}{\gamma+1} \left( 1-\frac{1}{Ma_1^2}\right)\right]^{\frac{-\gamma}{\gamma-1}} 
\end{align*}
As $s$ increases, $u$ decreases. $Ma_2$ is always $<1$, when $Ma_1 \rightarrow \infty$:
\begin{equation*}
    Ma_2 \rightarrow \sqrt{\frac{\gamma-1}{2\gamma}} = 0.38 ~ (\gamma = 1.4)
\end{equation*}
\begin{equation*}
    Ma_2^2 = \left( \frac{u_2}{a_2}\right)^2 = \left( \frac{u_2}{u_1}\right)^2 \left( \frac{u_1}{a_1}\right)^2 \left( \frac{a_1}{a_2}\right)^2 = \left( \frac{u_2}{u_1}\right)^2 Ma_1^2 \left( \frac{T_1}{T_2}\right)
\end{equation*}
\begin{equation*}
    Ma_2 = \sqrt{\frac{1 + \frac{\gamma -1}{\gamma + 1} \left( Ma_1^2 -1\right)}{1 + \frac{2\gamma}{\gamma + 1}\left( Ma_1^2 - 1\right)}}
\end{equation*}

A weak shock occurs at $Ma_1$ close to one. See page 31 for equation \\

\underline{\textbf{Rankine Hugoniot (p.32) - Adiabatic Shock (no $Ma$ dependency)}}
\begin{equation*}
    \frac{p_2}{p_1} = 1 + \frac{2 \gamma \left( \frac{\rho_2}{\rho_1}-1\right)}{\gamma + 1 - (\gamma -1)\frac{\rho_2}{\rho_1}}
\end{equation*} 
\underline{\textbf{Moving Shock Wave (p.33)}}

Switch to reference frame (from frame fixed with moving shock front into a frame moving with shock)
\begin{equation*}
    u_1 \widehat{=} u_s~,~ p_1 \widehat{=} p_0~,~ \rho_1 \widehat{=} \rho_0
\end{equation*}
Flow behind
\begin{equation*}
    u_2 \widehat{=} u_s-u_d~,~ p_2 \widehat{=} p_d~,~ \rho_2 \widehat{=} \rho_d
\end{equation*}
Shock $u_d$
\begin{equation*}
    u_d = u_s - u_2 = u_1 - u_2 = u_1 \left( 1 - \frac{u_2}{u_1}\right) = u_1 \frac{2}{\gamma+1}\left(1- \frac{1}{Ma_1^2}\right)
\end{equation*}
\begin{equation*}
    Ma_d = \frac{u_d}{a_d} = \frac{u_1-u_2}{a_d} = \frac{u_1}{a_1}\frac{a_1}{a_d}\left( 1- \frac{u_2}{u_1}\right) = Ma_1 \sqrt{\frac{T_1}{T_2}}\left(1-\frac{u_2}{u_1}\right)
\end{equation*}
\begin{equation*}
    u_d = \frac{a_0}{\gamma} \frac{\frac{\Delta p}{p_0}}{\sqrt{1 + \frac{\gamma+1}{2\gamma} \frac{\Delta p}{p_0}}}~~(a_1 \hat{=} a_0), ~~ Ma_s = \frac{u_s}{a_0} = \sqrt{1 + \frac{\gamma+1}{2\gamma}\frac{\Delta p}{p_0}}
\end{equation*}

Pressure increase
\begin{equation*}
    \frac{\Delta p}{p_0} = \frac{p_d - p_0}{p_0} = \frac{2 \gamma}{\gamma + 1} \left( Ma_S^2 -1 \right),~~[Ma_1 = \frac{u_1}{a_1} = \frac{u_s}{a_s}=Ma_s]
\end{equation*}

The ratio (Pressure increase) has an asymptotic limit. For high $Ma_s$, the function becomes limited. $\frac{u_s}{u_d} \rightarrow \frac{\gamma+1}{2}$ (for high pressure differences) \newline

\underline{\textbf{Detonations ($Ma_2 >1$) and Deflagrations ($Ma_2 < 1$) (p.36, ZND)}}

\textbf{Assumption: Ignore adiabatic flow, include however heat release}

Rayleigh line: $\frac{p_1}{p_0} = 1 + \frac{\rho_0}{p_0} u_0^2 - \frac{\rho_0}{p_0}\frac{\rho_1}{\rho_0}u_1^2 = 1 + \gamma Ma_0^2 \left( 1 - \frac{\rho_0}{\rho_1}\right)$,
Rankine Hugeniot with heat: $\frac{p_2}{p_0} = \frac{(\gamma+1)-(\gamma-1)\frac{\rho_0}{\rho_2}+2\gamma \hat{q}}{(\gamma+1)\frac{\rho_0}{\rho_2}-(\gamma-1)}~,~ \hat{q} = \frac{q_{heat}}{c_p T_1}$, This gives us $p_1$ and $p_2$, the pressure of the shockwave before the combustion and downstream after the combustion layer \\

\underline{\textbf{Chapman-Jouget Point (p.37)}}

...is the intersection where $Ma=1$, so $Ma_2 = 1 = Ma_0 \sqrt{\frac{\rho_0}{\rho_2}} \sqrt{\frac{\rho_0}{\rho_2}}$
The limiting case for shock cycle (Rayleigh tangent to Hugoniot Line ): 
\begin{equation*}
    \frac{\rho_0}{\rho_2} |_c = \frac{u_2}{u_0} |_c = \frac{ \gamma Ma_0^2 + 1 }{Ma_0^2 (\gamma + 1)}
\end{equation*}

Behind the shock, the flow is subsonic $\leftrightarrow$ strong detonation. There is a weak deflagration if the density ratio $\frac{\rho_1}{\rho_2} >> 1$. The reaction front propagates at subsonic speed.  Weak detonation: flow remains supersonic (not explainable through ZND)\\

\underline{\textbf{Laval Nozzle (p. 39)}}

Varying cross-section:
\begin{align*}
    \frac{p(x)}{p_0} &= \left[ 1 + \frac{\gamma - 1}{2} Ma^2 (x) \right]^{\frac{-\gamma}{\gamma-1}} \\
    \frac{A^*}{A(x)} &= Ma(x) \left[ \frac{2}{\gamma+1} + \frac{\gamma-1}{\gamma+1}Ma^2(x)\right] \\
    u(x) &= Ma(x) a_0 \frac{a(x)}{a_0} = Ma(x) a_0 \sqrt{\frac{T(x)}{T_0}} = \frac{a_0 \cdot Ma(x)}{\sqrt{1 + \frac{\gamma-1}{2}Ma^2(x)}} \\
    u^* &= a^* ~,~ \text{if}~ Ma^* = 1
\end{align*}

In order to increase the $Ma_{exit}$, reduce the area ration (tune $A^*$). Different flow regimes are shown on p. 41. A variable exit area is in practice not possible

\section{Unsteady one-dimensional Flows}

\textbf{Wave equation for small perturbations}
Assuming small pertubations around equilibrium state with first order pertubations will result into following differential equation (enthalpy):

\begin{align*}
    \frac{\partial p'}{\partial t} - a_0^2 \frac{\partial \rho'}{\partial t} &= 0 \Longleftrightarrow p' = a_0^2 \rho' \\
    \frac{\partial \rho'}{\partial t} + \rho_0 \frac{\partial u'}{\partial x} &= 0 ~~ (\text{mass eq.}) \\
    \frac{\partial u'}{\partial t} + \frac{a_0^2}{\rho_0} \frac{\partial \rho'}{\partial x} &= 0 ~~ (\text{momentum eq.})
\end{align*}

Through cross-differentiation (elimination of terms), one arrives at the d'Alembert solution:
\begin{align*}
    u'(x,t) &= a_0 [F(x-a_0t) + G(x+a_0t)] \\
    \rho'(x,t) &= \rho_0 [F(x-a_0t) + G(x+a_0t)]
\end{align*}
Through characteristics one defines left and right propagatiting waves, $F(\eta)$ and $G(\xi)$. The characteristics are in this case straight lines. Initial conditions are at $t=0$, boundary conditions are at $x=b.c.$ \\

\textbf{Method of characteristics for nonlinear wave propagation}
Here, no small pertubations are assumed, while assuming homentropic flow ($s=const.$). The Riemann invariants (characteristics) are not straight anymore, and can be curved. Disturbances are no longer constant, but have a flow dependent value. Given $a$ and $u$ are given along a curve $C$, find where it intersects with two characteristics, which cross at point $Q$. (See p. 48) \\ 

\textbf{Piston Motion in tube (example for unsteady one-dimensional motion)}:

\begin{itemize}
    \item Boundary Condition: At $x = x_p(t)$, $u(x=x_p,t) = u_p(t)$
    \item How to solve: Left propagating wave from rest state, intersects $P$ at $u=u_p$. The characterisitic with $\eta = const$ which then can intersect the other characteristic with $\xi = const. $ yields point $Q$
    \item $x = \left[ a_0 + \frac{\gamma + 1}{2} u_p(\tau)\right](t-\tau) + x_p(\tau)$
\end{itemize}

\textbf{Simple expansion waves} \\
In the case for the piston moving to the left, the characteristics are limited by two factors:

\begin{itemize}
    \item $x=a_0 t$: Initially, at $t=0$, the characteristic is maximum and can only be as steep as $a_0$
    \item $u_p=-U$: The piston motion can only have a max. velocity at its endpoints ($x_p = -Ut$ and $Ut$)
    \item This gives an area of solutions, which is called a ``centered fan''
\end{itemize}
\begin{align*}
    Ma &= \frac{|U|}{a_0} \left[ 1- \frac{\gamma -1}{2} \frac{|U|}{a_0} \right]^{-1},
    \frac{\rho}{\rho_0} = \left[ 1- \frac{\gamma-1}{2} \frac{|U|}{a_0} \right]^{\frac{2}{\gamma-1}} \\
    \frac{p}{p_0} &= \left[ 1- \frac{\gamma-1}{2} \frac{|U|}{a_0} \right]^{\frac{2 \gamma}{\gamma-1}}
\end{align*}

\textbf{Simple Compression Waves},
see p. 54, explained for increasing velocity to the right \\ \\ 
\underline{\textbf{Reflections}} \\
\textbf{Reflection from solid wall}: $G=-F$ if boundary moves with velocity $0$ \\

\textbf{Reflection from free boundary (contact surface), p.56}: The ratio $\alpha$ is the impendance, and is the ratio of both $a$ of two regions \\

\textbf{Reflection from an open end with outflow, p.58}: At an orifice ($a$ = outer, $0$ = stagnation), the characteristics are:

\begin{equation*}
    G = F - \frac{4}{\gamma-1} a(p_a)
\end{equation*}

The speed of sound is computed via the isentropic relations:

\begin{equation*}
    \frac{a_a}{a_0} = \sqrt{\frac{T_a}{T_0}} = \left( \frac{p_a}{p_0} \right)^{\frac{\gamma-1}{2\gamma}}
\end{equation*}
\section{Two-dimensional steady supersonic Flow}
An oblique shock wave forms around a body with a sharp tip or a long a wall with a sudden profile change (at the turning point). Two parameters describe the problem. The inflow Mach number $Ma_1$ and turning angle $\theta$. The velocity can be described by a component normal to the shock front and tangential to the shock front $u_n$ and $u_t$. The downstream and upstream pressures $p_1$ and $p_2$ are assumed constant.
\begin{align*}
    u_{2t} &= u_{1t} = u_t \\
    u_{1n} &= u_1 \sin \beta~,~u_{2n} = u_2 \sin(\beta-\theta) \\
    \frac{u_{2n}}{u_{1n}} &= \frac{(\gamma-1) Ma_1^2 \sin^2 \beta + 2}{(\gamma+1) Ma_1^2 \sin^2 \beta}\\
    \frac{u_{2t}}{u_{1t}} &= \frac{\tan (\beta-\theta)}{\tan \beta} = \frac{1-\tan\theta \cot \beta}{1 + \tan \theta \tan \beta} = \frac{1-(\tan\theta/\tan \beta)}{1 + \tan \theta \tan \beta} \\
    \tan \theta &= \frac{1-\frac{u_{2n}}{u_{1n}}}{\cot \beta + (\frac{u_{2n}}{u_{1n}})\tan \beta} = \frac{\left( \sin^2\beta - \frac{1}{Ma_1^2}\right)\sqrt{ 1- \sin^2 \beta}}{\sin \beta \left( \frac{\gamma+1}{2}-\sin^2 \beta + \frac{1}{Ma_1^2}\right)}
\end{align*}
The equation for $\theta$ is an implicit function for $\beta$, whereas there are two solutions for it. Normally, $\theta$ (geometry) and incident $Ma_1$ are given. For $Ma_1 > 1$, $\beta$ is in the range of $[\arcsin (1/Ma_1), 90^{\circ}]$. For a given $Ma_1$, there is a maximum turning angle $\theta_{max}$ beyond which the shock is not anymore at the turning point. The shock is then ``detached'' (set $Ma \rightarrow \infty$ in equation for $\theta$).
\begin{itemize}
    \item Strong and weak shock: The maxima of all shock angle for oblique shocks can be connected. Higher Mach angles $\beta$ occur with strong shocks
    \item Separation between $Ma_2 < 1$ and $Ma_2 > 1$
    \item For the changes in pressure, density and other thermodynamic properties, the equations for the normal shock can be used with a new ''effective Mach number'' $Ma_{1,new} = Ma_1 \sin \beta$
\end{itemize}
\begin{align*}
    u_{2n} &= u_2 \sin (\beta - \theta) \\
    Ma_2^2 \sin^2 (\beta - \theta) &= \frac{1 + \frac{\gamma - 1}{\gamma + 1} (Ma_1^2 \sin^2 \beta - 1)}{1 + \frac{2 \gamma}{\gamma + 1} (Ma_1^2 \sin^2 \beta -1)}
\end{align*}
\begin{itemize}
    \item Small turning angles $\theta$: In the limit of $\theta \rightarrow 0$, $\beta$ becomes the Mach angle $\mu = \lim_{\theta \rightarrow 0} \beta = \arcsin \frac{1}{Ma_1}$
    \item Hypersonic flow $Ma_1 >> 1$: Shock angle and turning angle linearly dependent on gas property $\gamma$ alone, $\sin \beta \approx \beta = \frac{\gamma + 1}{2} \theta$
\end{itemize}
\underline{\textbf{Continuous turning of supersonic flows}} \\
When the surface is moving continuously, the model can be adapted.
\begin{align*}
    \frac{dp}{p} &= - \frac{\gamma Ma}{1 + \frac{\gamma-1}{2}Ma^2} dMa \\
    d\theta &= - \frac{\sqrt{Ma^2-1}}{1 + \frac{\gamma-1}{2}Ma^2}\frac{dMa}{Ma}
\end{align*}

These equations describe the \underline{``Prandtl-Meyer Compression''}. In case the flow direction changes positively ($d\theta > 0$), the Mach number decreases ($dMa < 0$), and the pressure as well ($dp < 0$). 
\begin{align*}
\nu(Ma) &= \sqrt{\frac{\gamma + 1}{\gamma -1}} \arctan \sqrt{\frac{\gamma-1}{\gamma+1}(Ma^2-1)}-\arctan\sqrt{\text{Ma}^2-1}   \\
\nu_{max} &= \nu(Ma \rightarrow \infty) = \frac{\pi}{2}\left( \sqrt{\frac{\gamma+1}{\gamma-1}-1} \right) \hat{=} 130.5^{\circ} ~ (\gamma = 1.4)
\end{align*}

The above \underline{Prandtl-Meyer function} outputs degrees, which couple supersonic turns and change in Mach numbers.
\begin{equation*}
    \Delta \theta = \theta_2 - \theta_1 = -\nu(Ma_2) + \nu(Ma_1)
\end{equation*}
An oblique schock occurs at a distanced point $P$ away from the wall. A \underline{``Prandtl-Meyer Expansion''} occurse if the turning angle is negative. The flow accelerates, and Mach lines diverge. If the flow turns around a point, all lines are going along a center. \\ \\
\underline{\textbf{Reflection and crossing of waves}} \\
Use same equations as before. Here, following things must be considered:
\begin{itemize}
    \item Wedge with bounding wall: Reflections back from wall create additional shocks until $\theta < \theta_{max}$ (function of $\beta$). The Mach numbers decrease
    \item Wedge without bounding wall: Point $P$ is where fluid expands (reflected by the free jet boundary)
    \item Walls which converge: For two different turning angles $\theta$, two oblique shocks meet at point $P$:
    \begin{itemize}
        \item $\theta_1-\theta_5 = \theta_4 - \theta_2$
        \item $p_4(Ma_2, \theta_4, p_2) = p_5(Ma_3, \theta_5, p_3)$
    \end{itemize}
    \item Prismatic wing: First a oblique shock, then fans, then shock due to turning back flow to inflow direction. Even without friction, a drag is introduced
\end{itemize}
\underline{\textbf{Detached Shocks}}
Normally, the oblique shock occurs for $\theta < \theta_{max}$. For larger angles, the shock detaches and shows a near hyperbolic shape. See p.78 for details when it comes to decreasing $Ma$. \\

\underline{\textbf{Supersonic nozzle exit flows}} \\
Supersonic nozzle flow into stagnant environment.
\begin{itemize}
    \item Under-expanded scenario ($p_{env}<p_{noz}$): First PM-expansion due to reduction of pressure. However later on again goes back (rocket plume). Interacting incoming and reflecting expansion fans create the going back.
    \item Over-expanded scenario ($p_{env}>p_{noz}$): The turning angle $\beta$ is much smaller, the plume is impinged
\end{itemize}

\section{Method Characteristics for planar homentropic supersonic Flows}
For nonlinear wave propagation in ($x,t$)-space, one can compute flow fields in steady, two-dimensional homentropic flows. With the Crocco theorem, and the assumption of steady, inviscid, isoenergetic, adiabit flow, there is no vorticity. Introducing the velocity potential, we receive equations for the speed of sound. \\

\textbf{Transformation of equations (p.81)}\\
Mach lines are no longer straight lines over an extended region, but are treated as characterisitic curves on which the Riemann invariants remain constant. The characteristics become dependent on the local Mach angle $\mu = \mu(\nu)$ and the Mach number $Ma$, determined by the Prandtl-Meyer Function. The velocity magnitute is then given by the thermodynamic state ${p, \rho, T}$
\begin{equation*}
    F = \nu + \theta = const.~ \text{and}~ G = \nu - \theta = const.
\end{equation*} 
\textbf{Initial and Boundary Value Problems (p.85)}
\begin{itemize}
    \item Initial Value Problem: A curve C is given, as well as the Riemann Invariants. This gives us both Mach numbers and their flow directions in two points $1$ and $2$. In point $3$, both intersect, and this gives the local $\nu_3$ and $\theta_3$, meaning the local Mach number:
    \begin{align*}
        u_3 &= Ma_3 a_3 = Ma_3 a_0 \frac{a_3}{a_0} = Ma_3 a_0 \sqrt{\frac{T_3}{T_0}} \\
        a_0 &= \sqrt{\gamma R T_0},~~\frac{T_3}{T_0} = \left( 1 + \frac{\gamma-1}{2}Ma_3^2\right)^{-1}
    \end{align*}
\end{itemize}

\section{Homentropic Flow around slender Wings}
Instead of using methods of characteristics, now a new methodoligy is used. It uses also linearisations, which make the problems easier to handle and is valid for supersonic and subsonic flows. \\

\textbf{Initial Considerations}: \\
A function describes the wing profile with a camber line ($y_c(x)$) and a thickness distribution ($h(x)$), both depending on x. The boundary condition is that the flows are tangential to the contour, and that in the far field the stream velocity is free stream velocity $u_\infty$. \\

To simplify the problems, the coordinates are simplified as follows:
\begin{align*}
   & \tilde{x} = \frac{x}{L},~ \tilde{y} = \frac{y}{L},~ \sigma = \frac{c_{max}}{L},~\tau = \frac{h_{max}}{L},~\tilde{h}(\tilde{x}) = \frac{1}{h_{max}} h\left( \frac{x}{L} \right)\\
   & \frac{y_p}{L} = \sigma\tilde{y}_C(\tilde{x}) \pm \frac{\tau}{2} \tilde{h}(\tilde{x}),~\tilde{y}_C(\tilde{x}) = \frac{1}{c_{max}} y_C \left( \frac{x}{L} \right)
\end{align*}

Solving the equations for $c_p$ is not trivial, as they are non-linear. The parameters, on which the solution depends on are:
\begin{itemize}
    \item Thickness ratio $\tau$
    \item Camber ratio $\sigma$
    \item Angle of attack $\alpha$
    \item Free stream Mach number $Ma_\infty$
    \item Ratio of specific heats $\gamma$
\end{itemize}
\textbf{Linearized Theory} \\
The equations are simplified, assuming slender profiles ($\tau$, $\sigma$, $\alpha$ all $\ll 1$). This implies $h$ and $y_c$ remain small. Also, we assume small angles: $\alpha \approx \sin \alpha$
\begin{align*}
    c_p &= \frac{2}{\gamma Ma_\infty^2} \left\{ \left[ 1 - (\gamma-1)Ma_\infty^2 \tilde{u}_x \right]^{\frac{\gamma-1}{\gamma}}-1 \right\} \\
    & = -2 \tilde{u}_x(\tilde{x},\tilde{y}) = -2 \frac{\partial\phi}{\partial\tilde{x}}
\end{align*}

The pressure distribution is evaluated along the body surface, which in the linearised framework is $\tilde{y} = 0$ \\

\underline{\textbf{Linearised subsonic flows ($Ma\infty < 1$)}} \\
With the so-called \textbf{Prandtl Factor}, one can find a simple Laplace equation in the scaled coordinates ($\tilde{x},m\tilde{y}$):
\begin{align*}
    m &= \sqrt{1-Ma_\infty^2} \\
&\frac{\partial^2 \phi}{\partial \tilde{x}^2} + \frac{\partial^2 \phi}{\partial(m \tilde{y})} = 0
\end{align*}
Solutions can be found by superposition of fundamental solutions:
\begin{itemize}
    \item Source/Sink of magnitude $Q$ in ($\xi,0$): \\ $\phi(\tilde{x},\tilde{y}) = Q \ln \sqrt{(\tilde{x}-\xi)^2 + m^2\tilde{y}^2}$
    \item Vortex with circulation $\Gamma$ in ($\xi,0$): \\ $\phi(\tilde{x},\tilde{y}) = \Gamma \arctan \left( \frac{m \tilde{y}}{\tilde{x}-\xi} \right)$
\end{itemize}
\underline{\textbf{Symmetric profile w/o angle of attack ($\alpha =0$, $\sigma=0$)}} \\
The source $q$ chosen here is distribution per unit length, which can be defined as 
\begin{align*}
    \lim_{\tilde{y} \rightarrow \pm 0} \frac{\partial \phi}{\partial \tilde{y}} &= \pm q(\tilde{x}) = \pm \frac{\tau}{2}\frac{d \tilde{h}(\tilde{x})}{d \tilde{x}} = \pm \frac{\tau}{2} \tilde{h}'(\tilde{x}) \\
    \tilde{u}_x (\tilde{x},\tilde{y}) &= \frac{\tau}{2 m \pi} \int_0^1 \frac{\tilde{h}'(\xi)(\tilde{x}-\xi)}{(\tilde{x}-\xi)^2 + m^2\tilde{y}^2}d\xi \\
    \tilde{u}_y (\tilde{x},\tilde{y}) &= \frac{\tau}{2 \pi} \int_0^1 \frac{\tilde{h}'(\xi)m \tilde{y}}{(\tilde{x}-\xi)^2 + m^2\tilde{y}^2}d\xi \\
    c_p(\tilde{x}) &= -2 \tilde{u}_x (\tilde{x},\pm 0) = \\
    &= -\frac{\tau}{m\pi} \lim_{\varepsilon \rightarrow 0} \left[ \int_0^{\tilde{x}-\varepsilon} \frac{\tilde{h}'(\xi)}{\tilde{x}-\xi} d\xi + \int_{\tilde{x}+\varepsilon}^{1} \frac{\tilde{h}'(\xi)}{\tilde{x}-\xi} d\xi \right] \\
\end{align*} In the far field, the specific shape of the profile has no influence on the velocity components in $x$ and $y$. See eq. 10.37 and 10.40 for far-field behaviour.\\

\underline{\textbf{Cambered and flat plate in pitched flow (p.102) ($\tau = 0$, $\sigma \neq 0$, $\alpha \neq 0$)}} \\
A think plate which is cambered and a circulation term $\gamma$ with a circulation per unit length is chosen here:
\begin{align*}
    \gamma(\tilde{x}) = \frac{1}{\sqrt{\tilde{x}(1-\tilde{x})}} \left\{ C_1 - \frac{2}{m\pi}\int_0^1 \left[ \theta_P(\xi)-\alpha \right]\frac{\sqrt{\xi(1-\xi)}}{\tilde{x}-\xi} d\xi \right\}
\end{align*}
The so-called \textbf{Betz-Integral} describes this distribution. For flat plates with $\Theta_P = 0$, the equations simplify. The \textbf{Kutta Condition} postulates, that $C_1$ is chosen such that $u_x$ remains finite (at the end of the wing or profile).
\begin{align*}
    \gamma(\tilde{x}) &= -\frac{2\alpha}{m}\sqrt{\frac{1-\tilde{x}}{\tilde{x}}}~,~C_1 = -\alpha / m \\
    \tilde{u}_x(\tilde{x}, \pm 0) &= \pm \frac{\alpha}{m}\sqrt{\frac{1-\tilde{x}}{\tilde{x}}}~,c_p(\tilde{x}, \pm 0) = -2\tilde{u}_x(\tilde{x},\pm 0) = \pm \gamma(\tilde{x})
\end{align*}
In the far field, the velocity components are however also depending on $\alpha$. See equations (10.53) \\

\underline{\textbf{Thin plate with camber (p.106)}}: Pure camber problem, $c_p$ only depending on $\sigma$ \\

\underline{\textbf{Linearised superonic flows (p.108)}} \\
A new scaling parameter is introduced. Instead of the Prandtl Factor, one uses the \textbf{Scaling Parameter} $\lambda$. The equation for the wave equation is now hyperbolic. Similar to the linear, one-dimensional wave equation, a general solution can be defined in the form of two invariants.
\begin{align*}
    \lambda &= \sqrt{Ma_\infty^2-1} = \frac{1}{\tan(\mu_\infty)},~\frac{\partial^2 \phi}{\partial \tilde{x}^2} - \frac{\partial^2 \phi}{\partial(\lambda \tilde{y})^2} = 0 \\
    \phi(\tilde{x},\tilde{y}) &= F(\tilde{x}-\lambda y) + G(\tilde{x}+\lambda \tilde{y}) = F(\xi) + G(\eta)
\end{align*}
\underline{\textbf{Symmetric profile w/o angle of attack (p.109)}}
The solutions are deteremined by the profile slope $\tau \tilde{h}'(\tilde{x}) \approx \theta(\tilde{x})$. Along the surface
\begin{align*}
    &\frac{|u|}{u_\infty} \approx 1 + \tilde{u}_x(\tilde{x},0) = 1-\frac{\tau \tilde{h}'(\tilde{x})}{2\sqrt{Ma_\infty^2-1}} = 1 - \frac{\theta(\tilde{x})}{\sqrt{Ma_\infty^2-1}}\\
    &\frac{d}{d\tilde{x}}\left( \frac{|u|}{u_\infty} \right) = - \frac{\tau}{2\lambda} \tilde{h}''(\tilde{x}) \geq 0
\end{align*}
The flow is always accelerated after the shock ($|u| < u_\infty$) and reaches $u_\infty$ at the highest point of the profile ($\tilde{h}' = 0$) and is subsequently again accelerated until the trailing mach line at the end tip.
\begin{align*}
    \tilde{u}_x = \mp \frac{\tilde{u}_y}{\sqrt{Ma_\infty^2-1}}~\textbf{(Ackeret), } c_p(\tilde{x}) = \frac{\tau}{\sqrt{Ma_\infty^2-1}}\tilde{h}'(\tilde{x})
\end{align*}
As the $c_p$ increases (overpressure) and then decreases (underpressure) due to the change in profile height, appearance of \textbf{wave drag} can be explained.\\

\underline{\textbf{Cambered plate with angle of attack (p.111)}} \\
With a vanishing thickness $\tau = 0$, following equations can be computed:
\begin{align*}
    c_p(\tilde{x}) = -2\tilde{u}_x(\tilde{x}, \pm 0) = \mp \frac{2}{\sqrt{Ma_\infty^2-1}}\left[ \alpha - \theta_P(\tilde{x}) \right]
\end{align*}
Images show, that only for an inclination angle $\alpha$, lift can be achieved, independent of a camber with angle $\theta$\\

\underline{\textbf{Lift and drag (p.114)}}
\begin{align*}
    &C_L = C_{py} \cos \alpha - C_{px} \sin \alpha \approx C_{py} - \alpha C_{px} \approx C_{py} \\
    &\approx \int_0^1 \left[c_p(\tilde{x},-0)-c_p(\tilde{x},+0)\right] d\tilde{x} = \frac{F_y}{\frac{\rho_\infty}{2} u_\infty^2 Lb} \\
    &C_D = C_{py} \sin \alpha + C_{px} \cos \alpha \approx \alpha C_{py} + C_{px} \approx C_{px} \\
    &\approx \int_0^1 \left[  c_p(\tilde{x},+0)\left( \sigma \tilde{y}'_C + \frac{\tau}{2} \tilde{h}' \right) - c_p(\tilde{x},-0)\left( \sigma\tilde{y}'-\frac{\tau}{2}\tilde{h}' \right)\right] d\tilde{x}
\end{align*}
For the problems, following simplifications can be made:
\begin{itemize}
    \item Symmetric profile w/o angle of attack: $C_L = 0$, $C_D = \tau\int_0^1 c_p(\tilde{x})\tilde{h}'(\tilde{x}) d\tilde{x}$ 
    \item Flat plate with angle of attack: $C_L = 2 \int c_p(\tilde{x},-0)d\tilde{x}$. For subsonic, there is no drag (D'Alembert). For supersonic, there is a wave drag.
    \item Camber problem without pitch and thickness: $C_L$ same as for flat plate with angle of attack, but no Lift total.
\end{itemize}
\section{Homentropic Flow around axisymmetric slender Bodies}
Axisymmetric bodies with radial components are studied here with axial, radial, and azimuthal (in axisymmetric problems, the azimuthal component is not effecting the flow) components. To simplify the problems, the coordinates are defines as follows:
\begin{align*}
    &\Phi(x,r) = u_\infty L\left[ \tilde{x} + \phi(\tilde{x},\tilde{r}) \right]~~,~~ \tilde{x} = \frac{x}{L}~,~ \tilde{r} = \frac{r}{L} \\
    &u_x = \frac{\partial \Phi}{\partial x}~~,~~u_r = \frac{\partial \Phi }{\partial r}~~,~~\tilde{R}(\tilde{x}) = \frac{R(x)}{L}~~,~~A(x) = \pi R(x)^2 \\
    &c_p = \frac{p-p_\infty}{\frac{\rho_\infty}{2}u_\infty^2} = -2\tilde{u}_x - \tilde{u}_r^2
\end{align*}
\underline{\textbf{Linearised axisymmetric subsonic flow (p.123)}}\\
\begin{align*}
    &m^2 = 1 - Ma_\infty^2 \\
    &\phi(\tilde{x},\tilde{r}) = -\frac{1}{4\pi}\int_{0}^1 \frac{\tilde{A}'(\xi)}{\sqrt{(\tilde{x}-\xi)^2 + m^2 \tilde{r}^2}}d\xi \\
    &\tilde{A} = \pi \tilde{R}^2~~,~~\tilde{A}'' = 2 \pi \left[\tilde{R}'^2 + \tilde{R}\tilde{R}''\right] \\
    &\tilde{u}_x(\tilde{x},\tilde{R}) = \left[ \tilde{R}'^2 + \tilde{R}\tilde{R}'' \right] \ln \left( \frac{m\tilde{R}}{2\sqrt{\tilde{x}(1-\tilde{x})}}\right)~\textbf{Along surface $R$}\\ 
    & -\frac{1}{4\pi}\int \frac{\tilde{A}''(\xi)-\tilde{A}''(\xi)}{|\xi-\tilde{x}|} d\xi \\
    &\tilde{u}_r(\tilde{x},\tilde{R}) = \tilde{R}'(\tilde{x}) + \mathcal{O}(\tau^3)+\frac{m^2\tilde{R}}{4\pi}\int \frac{\tilde{A}'(\xi)-\tilde{A}'(x)}{\sqrt{(\tilde{x}-\xi)^2 + m^2\tilde{R}^2}}d\xi \\
    &c_p \approx -2\left[ \tilde{R}'^2 + \tilde{R}\tilde{R}'' \right] \ln \frac{\sqrt{1-Ma_\infty^2 }\tilde{R}}{2\sqrt{\tilde{x}(1-\tilde{x})}} \\
    &+\frac{1}{2\pi} \int_0^1 \frac{\tilde{A}''(\xi)-\tilde{A}''(x)}{|\xi-x|} d\xi - \tilde{R}'^2(\tilde{x}) \\
    &\tilde{u}_x(\tilde{x},\tilde{r}) = -\frac{1}{4\pi}\frac{2\tilde{x}^2-m^2\tilde{r}^2}{\sqrt{\tilde{x}^2+m^2\tilde{r}^2}}\int_0^1\tilde{A}(\xi)d\xi~~\textbf{far field} \\
    &\tilde{u}_x(0,\tilde{r})=\frac{1}{4\pi m^3\tilde{r}^3}\int_0^1 \tilde{A}(\xi)d\xi \propto \frac{1}{\sqrt[3]{1-Ma_\infty^2}\tilde{r}^3}
\end{align*}
The disturbance strength is proportional to the volume of the body in the flow, compared to the planar case. For the flow, one assumes a continous heat/sink flow from the front to the end tip \\

\underline{\textbf{Linearised axisymmetric supersonic flow (p.126)}} \\
Compared to before, one has hyperbolic differential equations with characterisitics $\tilde{x}\pm \lambda \tilde{r} = const.$
\begin{align*}
    &\lambda^2 = Ma_\infty^2-1 > 0 \\
    &\phi(\tilde{x},\tilde{r}) = \int_{arccosh(\tilde{x}/\lambda\tilde{r})}^0 q (\tilde{x}-\lambda\tilde{r}\cosh z)dz \\
    &\tilde{u}_x(\tilde{x},\tilde{R}) = \left[ \tilde{R}'^2 + \tilde{R}\tilde{R}'' \right] \ln\left[ \frac{\lambda \tilde{R}(\tilde{x})}{2\tilde{x}}\right]+\frac{1}{2\pi}\int_0^{\tilde{x}} \frac{\tilde{A}''(\tilde{x})-\tilde{A}''(\xi)}{\tilde{x}-\xi} d\xi \\
    &\tilde{u}_r(\tilde{x},\tilde{R}) = \frac{1}{2\pi}\int_0^{\tilde{x}-\lambda\tilde{R}} \frac{\tilde{A}''(\xi)(\tilde{x}-\xi)}{\sqrt{(\tilde{x}-\xi)^2-\lambda^2\tilde{R}^2}} d\xi \approx \tilde{R}'(\tilde{x}) \\
    &c_p(\tilde{x}) =\frac{\tilde{A}''(\tilde{x})}{\pi}\ln \frac{\sqrt{Ma_\infty^2-1}\tilde{R}}{2\tilde{x}} - \frac{1}{\pi}\int_0^{\tilde{x}} \frac{\tilde{A}''(\tilde{x})-\tilde{A}''(\xi)}{\tilde{x}-\xi} d\xi - \tilde{R}'^2 \\
    &c_D = \frac{1}{\pi R_{max}^2} \int_0^1 c_p(\tilde{x})\tilde{A}'(\tilde{x})d\tilde{x}
\end{align*}
\setcounter{section}{12}
\section{Steady flows with friction and heat transport}
\underline{\textbf{Two-dimensional Flows (p.141)}} \\
In two dimensional flows, the friction or inviscous flow effects can be seen in the momentum conservation and energy conservation laws, whereas the full stress tensor $\tau_{ij}$ is kept, leading to the Navier-Stokes equations. The viscous dissipation $\Phi$ is then also defined, together with the heat conduction term. \\

\underline{\textbf{Parallel Flows (p.142)}}\\
In parallel flows, one assumes only a velocity in x-direction ($u(x,y)$) \\

\underline{\textbf{Flow through normal shock (p.142)}}\\
A flow before (${u_1,T_1}$)and after (${u_2,T_2}$) the shock is considered. Compared to the shock without thickness, there is now a shock thickness $L$, which is finite. The flow properties are now changing smoothly from the upstream to the downstream values while maintaining local equilibrium. Before and after the shock, the jumb conditions apply. In the shock region with thickness $L$. This demonstrates that the entropy increase is linked to heat conduction and viscous dissipation. \\

\underline{\textbf{Channel and pipe flow with constant cross section (p.144)}}\\
The enthalpy equation is integrated. The temperature related term is reduced to a heat flux $\dot{Q}_{12}$. The integral basically becomes an averaging equation for different heat fluxes and friction forces. The forces with regards to the friction forces can also be reduced through the ``Average'' fo the momentum equation. \\

\underline{\textbf{Pipe flow with friction and heat input (p.145)}} \\
If heat fluxes through the wall and wall friction are condsidered, the differentialforms of the momentum and enthalpy equations are studied. Furthermore, a fricion coefficient $\lambda = \lambda(Re, \varepsilon/D)$ is introduced.This friction coefficient is only depending on the relative wall roughness $\varepsilon/D$ for turbulent flows with large Reynolds numbers. This yields a general formulation, which can be studied if certain parameters change or have specific values:
\begin{equation*}
    (1-Ma^2)\frac{du}{u} = \frac{\lambda}{2}\gamma Ma^2 \frac{dx}{D} + \frac{\gamma-1}{a^2} dq
\end{equation*}
\begin{itemize}
    \item Generally, in subsonic flow ($Ma < 1$), the velocity ($du > 0$) increases because of wall friction and in case of heat input ($dq > 0$) 
    \item In supersonic flow, the behavious is opposite, and $du < 0$ given a heat input
    \item If $du = 0$ is to be maintained, then $dq < 0$. It follows with ideal gas equations that the entropy will decrease $ds < 0$
    \item If the entropy shall remain constant ($ds = 0$), then the density and velocity do not remain constant but change in to isentropic flow: $(1-Ma^2)\frac{du}{u}=\lambda \frac{Ma^2}{2}\frac{dx}{D}$ 
    \item For an adiabatic wall $dq=0$, the entropy always increases, then the equation becomes: $\left( \frac{1}{Ma^2}-1 \right)\frac{du}{u} = \frac{\lambda}{2}\gamma \frac{dx}{D}$. This means, that for $Ma <1$, the velocity increases, and for $Ma > 1$, the velocity decreases. 
    \item The differential equation above can be solved by utilising the critical Mach number, where the flow velocity is exactly $Ma^*$. This yields an implicit equation which can give an indication at which propagation distance $x$, respectively $x_{max}$ the flow becomes sonic. In order to receive an explicit solution, a condition where $x=x_1$ and its associated $Ma_1^*$ must be known.
    \item Based on this, the temperature distribution along $x$ can be computed: $\frac{T(x)}{T^*} = \frac{T_1}{T^*} + \frac{\gamma-1}{2}\left[Ma_1^{*2}-Ma^{*2}\right]$. This shows, that temperatures decrease with $Ma_1^* < 1$, and increase with $Ma_1^* > 1$
\end{itemize}

\underline{\textbf{Planar Couette Flow (p.149)}} \\
In a pipe or planar channel with constant cross-section, there is always a change in flow properties with propagation distance $x$ due to friction. Choosing the appropriate heat removel can maintain velocity and density, but pressure and temperature would have to decrease in the flow direction.\\
A fully developed flow can be reached, though, if energy can be added to the flow in form of mechanical work. Then the flow and thermodynamic properties remain constant along $x$. An example for this is a moving upper wall above a lower, fixed wall. The solution for the temperature distribution as function of local velocity is ($Pr = \mu c_p / k$):

\begin{equation*}
    \frac{T(u)}{T_\infty} = \frac{T_w}{T_\infty}+ \left[1-\frac{T_W}{T_\infty}\right]\frac{u}{U_\infty} + Pr \frac{\gamma-1}{2}Ma^2_\infty \frac{u}{U_\infty}\left[1-\frac{u}{U_\infty}\right]
\end{equation*}

In order to comput the velocity distribution, there are two methods: One assumes that the viscosity is depending on temperature, the other one assumes constant viscosity:

\begin{itemize}
    \item $\mu = \mu(T) = \mu_\infty \frac{T(u)}{T_\infty}$, $\rho \propto \frac{1}{T}$:
    \begin{itemize}
        \item $\frac{\tau_w h}{U_\infty \mu_\infty} = \frac{1}{2}\left[ 1 + \frac{T_w}{T_\infty} \right] + \frac{1}{6}Pr \frac{\gamma-1}{2}Ma_\infty^2$
        \item $c_f = \frac{\tau_w}{\frac{1}{2}\rho_\infty U_\infty^2} = \frac{1}{Re}\left[  1 + \frac{T_w}{T_\infty} + \frac{1}{3}Pr \frac{\gamma-1}{2}Ma_\infty^2 \right]$ 
    \end{itemize}
    \item $\mu = \mu_\infty$:
    \begin{itemize}
        \item $\frac{u}{U_\infty} = \frac{y}{h}~~,~~ \frac{\tau_w h}{\mu_\infty U_\infty} = 1~~,~~ c_f = \frac{2}{Re}$
    \end{itemize}
\end{itemize}
The heat flows are computed as follows:
\begin{align*}
    &q_w (y=0) = -\frac{c_p T_\infty}{Pr U_\infty}\left[ 1 - \frac{T_w}{T_\infty} + Pr \frac{\gamma-1}{2} Ma_\infty^2 \right] \cdot \tau_w \\
    &q_\infty = k \frac{dT}{dy}\vert_{y=h} = \frac{c_p T_\infty}{Pr U_\infty}\left[ 1 - \frac{T_w}{T_\infty}-Pr \frac{\gamma-1}{2}Ma_\infty^2 \right]\cdot \tau_w \\
    &q = q_w + q_\infty = - \frac{c_p T_\infty}{Pr U_\infty}\left[ Pr (\gamma-1) Ma_\infty^2  \right]\cdot \tau_w = - U_\infty \tau_w
\end{align*}
The signs follow from the positive y-direction to from bottom to top. The last equation shows, that the heat flux must exaclty match the the amount of mechanical power introduced by wall motion ($U_\infty \tau_w$). If both walls were adiabatic, a fully developed flow state could not be maintained. \\

If the lower, fixed wall were adiabatic ($q_w = 0, q = q_\infty$), following relations would hold true:
\begin{align*}
    &\frac{T_w}{T_\infty}\vert_{q_w = 0} = \frac{T_r}{T_\infty} = 1 + Pr \frac{\gamma-1}{2}Ma_\infty^2, r = Pr = \frac{T_r-T\infty}{T_0-T_\infty} = Pr \\
    &\frac{T}{T_\infty} = 1 + Pr \frac{\gamma-1}{2}Ma_\infty^2\left[ 1- \left(  \frac{u}{U_\infty}\right)^2 \right]
\end{align*}
\underline{\textbf{Flow in laminar boundary layers (p.154)}} \\
The momentum layer thickness scales with $\delta / L \propto 1 / \sqrt{Re}$ and the thermal layer scales with $\delta_T / L\propto 1/\sqrt{Re Pr}$, whereas $Pr$ is close to 1. It is assumed that the pressure along the the layer in $y$ is constant, so $p(x)$. The boundary conditions are the no slip condition, adiabatic and the outer edge of the boundary layer have a specific velocity and temperature $u_{\delta}$, $T_\delta$.
\begin{equation*}
    \frac{T}{T_\delta} = 1 + \frac{\gamma-1}{2}Ma_\delta^2\left[ 1- \left( \frac{u}{u_\delta}\right)^2 \right]
\end{equation*} 
The recovery temperature is given when $u=0$ at the wall, which is identical to the stagnant temperature of the flow $T_0$. \\

\underline{\textbf{Wall conditions (p.156)}}\\
Along the wall, the differential equations give insights regarding the pressure and velocity profile. For an adiabatic wall, following realtions hold true:
\begin{align*}
    &\mu \frac{\partial^2 u}{\partial y^2}\vert_{y=0} = \frac{dp}{dx} \\
    & \mu \frac{\partial^2 u}{\partial y^2}\vert_{y=0} = - \frac{d \mu}{dT} \frac{\partial T}{\partial y} \frac{\partial u}{\partial y}\vert_{y=0} \\
    & \tau_w = 0 = \mu \partial u/\partial y |_{y=0} \text{  (State of flow separation)}
\end{align*}
The above equations show that the curvature of the velocity profile is determined by the outer flow pressure gradient. The flow separation is determined by the condition given. For flat plates, $dp/dx = 0$, hence the profile curvature is zero at an adiabatic wall. The general equation 2 shows temperature dependent viscosity. \\

In general, one can assume $d\mu / dT > 0$ for gases, and $\partial u / \partial y |_{y=0} > 0$ applies. This means, that if heat is removed (temperature gradient positive), that the curvature of the velocity profile is negative, and the flow accelerates ($dp/dx < 0$). For heating, the flow is decelerated. One can show that the temperature profile has always negative curvature ($\partial^2 T / \partial y^2 < 0$). For $Pr = 1$, the total enthalpy remains constant, $h_0 = h_{0\delta}$.

\underline{\textbf{Flat plate boundary layer (p.159)}}\\
Here, one differentiates between for $Pr = 1$ and $Pr \neq 1$:
\begin{itemize}
    \item $Pr = 1$: $c_f = \frac{\tau_w}{\frac{1}{2}\rho_\infty U_\infty^2}~~,~~c_q = \frac{q_w}{\rho_\infty U_\infty^3} = \frac{T_w-T_r}{T_\infty}\frac{c_f}{2(\gamma-1)Ma_\infty^2}$
    \item $Pr \neq 1$: Numerical solution, and model temperature with Sutherland formula
\end{itemize}
\underline{\textbf{Velocity and temperature profiles in a compressible flat boundary layer (p.161)}}\\
Calculate according to script


\end{multicols*}
\end{document}

